Settler colonial projects and variation in paths to building the welfare and carceral state

### 2023-11-21 t1_analysis.Rmd
ff

---
title: "Varying paths to building the social state in the progressive era"
author: "Frank Edwards"
date: "last updated 11/1/23"
output: bookdown::html_document2
---

```{r setup, include = F, message = F}
knitr::opts_chunk$set(message = F, echo = F, warning = F)
library(tidyverse)
library(lme4)
library(maps)
library(usmap)
library(grid)
library(gridExtra)
options(scipen = 999)
options(dplyr.print_max = 1e9)
map<-map_data("state")

@@

1. pull immig + urban data
2. pull tenant farmer data
3. model employment -> each, consider loglinear process?

theory re: white immigs: coercion / rehab in NE, yeoman uplift in W, frontier as panacea for urban, lots of hist on that, gets at SocWork/Police vs Teachers employment

@@

(GENERAL FRAMEWORK: Racial capitalism / settler colonization) MATERIAL CONDITIONS FOR SETTLER COLONIZATION STRUCTURE POLICY DESIGN. socio-political processes generate diverse white elite interests

SETTLER COLONIALISM AS STRUCTURE - PROCESS. BUT SPATIALLY VARIABLE DEPENDING ON LOCAL INTERESTS

1. Frontier settler polity -> elite interest in disposession / assimiatilion -> aggressive paternalistic welfare state for AIAN, social democracy for settlers
2. Jim Crow -> elite interest in exploitation and exclusion -> elite paternalism, malign neglect, hard coercion
3. Metropole settler polity -> elite interests in labor discipline / immig assimilation -> progressive paternalism, strong coercive welfare state for W, hard coercion and contested welfare for Blk

THESE ARE MULTIVARIATE INDICATORS, CAN USE CLUSTERING TO CHECK IT AND MEASURE IT.

1: MEASURE VARIATION IN CONTEXT, ASSUME CONTEXT PRODUCES INTERESTS. 2: MEASURE VARIATION IN POLICY DEVELOPMENT.

3 dimensions: frontier, antiBlack, white immig assim (urban)

frontier: {homestead claims, tribal density, https://www.census.gov/dataviz/visualizations/001/}
railrooad (corporate) + homestaed (individual) + morrill (state) - nice three piece approach, more complex and complete story.

antiblack: {jim crow polit econ (sharecropping), prop black}

urban assim: {urbanization, industrial employment?, urban white immig, log black pop}

@@

Interests of social policy actors defined within these fields. Bourdieusian historical stuff?

@@

interests can be observed indirectly sometimes, right? homesteading is direct evidence of interest. tenant farming (or land ownership) is direct evidence of interest. industrial urban is tougher, but maybe there we just go indirect and use urbanization + racial demography.

I THINK THIS WORKS. CLUSTER ON INTERESTS + DEMOGRAPHY. OUTCOMES ARE POLICY MEASURES: POLICY OUTCOMES (EG EMPLOYMENT, SPENDING, GROUP QUARTERS POPS, ENROLLMENTS).

TRY TO DO ONLY A FEW BIG NATIONAL PARAMETRIC TESTS. LOTS OF DESCRIPTIVE SUBSET WORK

TEST 1: clustering, spatial clustering

then TONS of descriptives to paint pictures of regional social / political contexts

TEST 2: regression on outcomes

@@

dat<-dat %>%
  left_join(hs)

# # on local read temp
# dat<-read_csv("./data/temp_data.csv")
# hs<-read_csv("./data/temp_hs.csv")
```

@@

@@

### STATES, TERRITORIES, AND NATIVE LANDS

@@

### DIRECT AND PROXY MEASURES FOR ELITE SETTLER INTERESTS

@@

### Measuring anti-Blackness and post-Reconstruction racial politics

POPULATION ALONE DESCRIPTIVES

@@

```{r}
dat %>% 
  filter(year == 1930) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = log(blk))) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis") + 
  theme_void() + 
  theme(legend.position = "bottom")
```

@@

@@

### Measuring the size, power, and interests of the expropriative settler polity

@@

3) PRESENCE OF NATIVE LAND HOLDINGS - VERY CLEAR

@@

```{r ts-homestead-claims, fig.cap = "Cumulative Homestead act claims by state and year"}
ggplot(hs %>% 
         filter(year < 1940),
       aes(x = year,
           y = hs_claims,
           group = state)) + 
  geom_line(alpha = 0.5) + 
  geom_text(data = hs %>% 
              filter(year == 1935),
            aes(x = year,
                y = hs_claims,
                label = state),
            position = position_nudge(x = 8),
            alpha = 0.7) + 
facet_wrap(~region)  + 
  labs(y = "Number of claims",
       x = "Year",
       subtitle = "") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

THIS IS COOL! Look at homestead (as settler polity) plus mode (as boarding school on / off)

LOOK AT ACRES PER CLAIM? GET AT BOTH POP SIZE AND RELATIVE RESOURCES?

```{r}
ggplot(hs %>% 
         filter(year == 1961) %>% 
         mutate(region = factor(region, 
                                levels = c("West", 
                                           "North Central",
                                           "South",
                                           "Northeast"))),
       aes(x = hs_acres / 1e6,
           y = fct_reorder(state, hs_acres))) + 
  geom_point() + 
  labs(x = "Acres claimed (millions)",
       y = "",
       color = "Census Region") + 
  facet_wrap(~region, 
             scales = "free_y")
```

pop / claims ratio is a little funky, but probably useful. can peg to 1940 census.

WOWZERS. 
20+pct of pop likely had HS claim land in 3 states.

Wow, let's pull through this. ONE INDICATOR OF SETTLER POLITY. DEmographic base is one dimension - cultures / institutions/ resources also key. but useful!

```{r}
dat %>% 
  filter(year == 1920) %>% 
  mutate(hs_pop = hs_claims / tot_pop) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = hs_pop)) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis") + 
  theme_void() + 
  theme(legend.position = "bottom")
```

### Measuring the magnitude of social problems Progressive reformers would seek to 'fix'

PULL IMMIG + URBAN - UNIQUE COMBO THAT PUSHES FORWARD WELFARE STATE

```{r immig-ts, fig.cap = "White immigrant population by state 1870-1940"}
p1<-ggplot(dat,
       aes(x = year, y = white_immig / tot_pop,
           group = state)) + 
  geom_line(alpha = 0.5) + 
  labs(subtitle = "Proportion white immigrant",
       y = "Proportion",
       x = "Year") + 
  facet_wrap(~region)

p2<-ggplot(dat%>% 
  filter(!(is.na(region))),
       aes(x = year, y = white_immig,
           group = state)) + 
  geom_line(alpha = 0.5) + 
  scale_y_log10() + 
  labs(subtitle = "White immigrant population",
       y = "Count (note log10 axis)",
       x = "Year") + 
  facet_wrap(~region)

Relative population share as first generation immigrant declines in NC across period uniformly. In NE, increase until 1920 for most states, then monotonic decline. Low levels in S. Declining levels in W.

Absolute white immigrant pop numbers increasing in NE, Increasing, then plateau in NC, increasing then plateau in W, and stable and low in S.

### urban

```{r urban-ts, fig.cap = "Urban white immigrant population by state 1870-1940"}
p1<-ggplot(dat%>% 
             filter(year<=1930) %>% 
             filter(!(is.na(region))),
           aes(x = year, y = urban_immig_pop / tot_pop,
               group = state)) + 
  geom_line(alpha = 0.5) + 
  labs(subtitle = "Proportion of population white immigrant urban",
       y = "Proportion",
       x = "Year") + 
  facet_wrap(~region)

p2<-ggplot(dat%>% 
             filter(year<=1930) %>% 
             filter(!(is.na(region))),
           aes(x = year, y = urban_immig_pop,
               group = state)) + 
  geom_line(alpha = 0.5) + 
  labs(subtitle = "White immigrant urban population",
       y = "Count",
       x = "Year")+ 
  facet_wrap(~region)

@@

```{r}
dat %>% 
  filter(year == 1920) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = log(white_immig))) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis") + 
  theme_void() + 
  theme(legend.position = "bottom")
```

# distributions of policy outcomes

### the distribution of social state capacity building

@@

#### kids in institutional group quarters

@@

```{r gq-box, fig.cap = "Children in institutional group quarters 1930"}
dat %>% 
  filter(year == 1930) %>% 
  mutate(across(child_inst_correctional:child_inst_poor_disabled, 
          ~ .x/children * 1e3)) %>% 
  pivot_longer(cols = child_inst_correctional:child_inst_poor_disabled,
               names_to = "type",
               values_to = "pop") %>% 
  ggplot(aes(x = type, y = pop, 
             color = region)) + 
  geom_boxplot()

@@

employment trajectories look kinda log-linear. almost certainly non-linear, so need a curve on trajectories regardless. log probably solves it well enough.

q1: salience of settler expropriotion -> state formation

four models, each employment outcome + children in poverty institutions

start with lm for log(rates)

trouble fitting models for sparse data (e.g. social workers, children in gq), but works fine and in expected directions for teachers, police.

switch to brms for estimation on tricky ones

```{r}
### FIT PROBLEM WAS WITH EARLY YEARS NEAR ZERO AND LOW VARIANCE
model_dat <- dat %>% 
  filter(year>=1880) %>% 
  mutate(police_rate = (police + 1) / tot_pop * 1e5,
         socwork_rate = (socwork + 1) / tot_pop * 1e5,
         teacher_rate = (teachers + 1) / tot_pop * 1e5,
         child_inst_rate = (child_inst_poor_disabled + 0.1) / 
           children * 1e5,
         prop_blk = (blk + 1) / tot_pop,
         hsclaim_prop = hs_claims / tot_pop,
         year = scale(year))

sc_police<-lmer(log(police_rate) ~ 
              scale(hsclaim_prop) +
              year + 
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))

sc_teacher<-lmer(log(teacher_rate) ~ 
              scale(hsclaim_prop) +
              year + 
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))

sc_socwork<-brm(log(socwork_rate) ~ 
              scale(hsclaim_prop) +
              year + 
              (1 + year|state),
            data = model_dat)

sc_gq<-brm(log(child_inst_rate) ~ 
              scale(hsclaim_prop) +
              year + 
              (1 + year|state),
            data = model_dat)

jc_police<-lmer(log(police_rate) ~ 
              scale(prop_blk) +
              year + 
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))

jc_teacher<-lmer(log(teacher_rate) ~ 
              scale(prop_blk) +
              year + 
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))

jc_socwork<-lmer(log(socwork_rate) ~ 
              scale(prop_blk) +
              year + 
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))

jc_gq<-lmer(log(child_inst_rate) ~ 
              scale(prop_blk) +
              year + 
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))
```

This all looks fine. How do I want to visualize this?

we'll eventually do trajectory visuals for each regime type - what is the expected trajectory under different regimes, where do we end up in 1930? Think about how to handle to intercepts / slopes. What would a zero mean? pulling out everything not directly in national trend and regime measures. is that what I'm after? I guess I really do just want the regime-specific trend lines, in which case E[rate | regime] should zero out the individual idioscyncracies. it's oversimplified, but that should be ok. Just be super clear on what the iterpretation is on prediction. it's not a hypothetical state, its the mean expectation for the slope of the regime var.

model is

\[\hat{Y} = \alpha + \beta \textrm{Regime} + \delta year + \epsilon\]

Here we are very much interested just in $\beta$, the rest are nuissance parameters. that works. can vis e[state type] under diffferent REGIME sets.

Q: are there places with competing regime sets? CA, WA, others? where we need a multivariate regime model?

```{r}
new_dat<-data.frame(year =  c(1.3, 1.3, 1.3),
                    hsclaim_prop = c(0, 0.05, 0.15))
t<-predict(sc_police, 
           newdata = new_dat,
           re.form = ~0,
           type = "link")

t1<-predict(sc_teacher, 
           newdata = new_dat,
           re.form = ~0,
           type = "link")

exp(t1)

>>>>

---
title: "Settler colonial projects and variation in paths to building the welfare and carceral state"
author: "Frank Edwards"
date: "last updated `r Sys.Date()`"
output: bookdown::html_document2
abstract: "Place and time variation in problems of settler colonization are related to the kinds of governments that local settler elites build. "
---

```{r setup, include = F, message = F}
knitr::opts_chunk$set(message = F, echo = F, warning = F)
library(tidyverse)
library(lme4)
library(maps)
library(usmap)
library(grid)
library(gridExtra)
library(brms)
options(scipen = 999)
options(dplyr.print_max = 1e9)
map<-map_data("state")

@@

leaving out railroad grants induces bias if morill or homestead grants are negatively correlated with railroad grants.

2. Lit review
3. writeup demographics and bivars
4. models

@@

HOW ARE PROJECTS / PROBLEMS OF SETTLER COLONIAL GOVERNANCE RESOLVED? WHAT ORGS AND INSTITUTIONS ARE BUILT?

Settler colonial regimes seek to resolve different problems across places and within places over time. This paper proposes a quantitative framework for understanding how the geographic process of settler colonization results in different styles of governance across places. We show that a series of historical economic and demographic measures (proxies for variation in modes of settler colonial extraction and exploitation) are closely related to trajectories of state formation across places, measured as employment levels in core social policy system sectors.

At a sub-national level, settler colonial regimes seek to address a range of problems inherent to their nation-building project: the dispossession of indigenous lands, the erasure of indigenous peoples, the exploitation of racially marginalized laborers, and the reproduction and expansion of the settler polity. We show that the historical geography of these problems is closely related to the policy infrastructure that governments develop. The consequences of this variation in trajectories are likely broad and varied.

Historians tell us these projects were active locally. 1. Frontier; 2. Jim Crow; 3. Metropole. Can we use demographic, legal, and employment data to map it out / model it?

@@

@@

<!-- 1. Frontier settler polity -> elite interest in disposession / assimiatilion -> aggressive paternalistic welfare state for AIAN, social democracy for settlers -->
<!-- 2. Jim Crow -> elite interest in exploitation and exclusion -> elite paternalism, malign neglect, hard coercion -->
<!-- 3. Metropole settler polity -> elite interests in labor discipline / immig assimilation -> progressive paternalism, strong coercive welfare state for W, hard coercion and contested welfare for Blk -->

IMPLICATIONS - a) settler colonization as structure / structuring: context is key, and when SC is active, likely can be predictive; b) durable consequences for policy institutions (legacies, etc); c) comparative state formation / comparative welfare / carceral states.

# What is already known about settler colonization and state formation in late 19th and Early 20th c.

## Frontier

review here

## Jim Crow

review here

## Metropole

review here

@@

dat<-dat %>%
  left_join(hs) %>% 
  filter(year>=1880, year<=1920)

# # on local read temp
# dat<-read_csv("./data/temp_data.csv")
# hs<-read_csv("./data/temp_hs.csv")

### land grab u data
# lgu_cessions<-read_csv("./data/landgrabu-data/Morrill_Act_of_1862_Indigenous_Land_Parcels_Database/CSVs/Cessions.csv")
### this is cessions from the Royce maps, overlaps with morrill, but includes prior cessions as well
### would need to use shapefiles to portion out by state
lgu_parcels<-read_csv("./data/landgrabu-data/Morrill_Act_of_1862_Indigenous_Land_Parcels_Database/CSVs/Parcels.csv")
lgu_unis<-read_csv("./data/landgrabu-data/Morrill_Act_of_1862_Indigenous_Land_Parcels_Database/CSVs/Universities.csv")
```

@@

We are generally interested in evaluating whether a set of policy outcomes {employment, infrastructure} are related to a set of demographic and economic predictors {population composition, economic activity, ownership regime}.

To do so, we rely on historical data from the US Census accessed via IPUMS (Ruggles et al), from the Bureau of Land Management (Archived by the University of Nebraska), from the US Department of the Interior (CITE BS REPORT), and from records our team digitized from the Commissioner on Indian Affairs (CITE).

Less interested in causal identification, more interested in understanding spatial patterns and trajectories. We already know there was spatial variation in settler colonial contexts and policy systems, trying to see if we can capture it systematically with national quant data. Move from specific to general.

demonstrate var(infrastructure), show e[infrastructure] = f(settler interests)

@@

@@

geography itself endogenous to settler colonization, this is tricky!

## sources of data

Census; ARCIA; Interior; etc

## statistical methods and models

EDA; multilevel models on outcome rates (state intercepts, state x year slopes, log outcomes for nonlinear process, etc) estimated with hmc, weak priors

# Findings

@@

## Geographic contexts for settler colonial policymaking across US states 1880 - 1920

### Jim Crow and the salience of anti-Blackness for local social and carceral policymaking

Figure \@ref(fig:blk-map) shows the distribution of the recorded Black population across US states in 1920. Panel 1 displays this data as the relative population composition of Black residents relative to to all residents, and panel 2 displays the absolute size of the Black population across US states.

DISCUSS THE DISTRIBUTION

```{r blk-map, fig.cap = "Black population, US States 1920 (note log scale)"}
p1<-dat %>% 
  filter(year == 1920) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = blk / tot_pop * 100)) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis") + 
  theme_void() + 
  theme(legend.position = "bottom") + 
  labs(fill = "Percent of total",
       subtitle = "Panel 1. Population composition")

p2<-dat %>% 
  filter(year == 1920) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = blk + 100)) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis", trans = "log10") + 
  theme_void() + 
  theme(legend.position = "bottom") + 
  labs(fill = "Population\n(note logarithmic scale)",
       subtitle = "Panel 2. Total population")

grid.arrange(p1, p2, nrow = 1)
```

@@

@@

```{r}
dat %>% 
  filter(year == 1920) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = blk_sharecrop + 10)) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis",
                        trans = "log10") + 
  theme_void() + 
  labs(fill = "Number of Black residents\non tenant farms\n(note logarithmic scale)")
```

@@

### Settlers and indigenous disposession

@@

```{r}
p1<-dat %>% 
  filter(year == 1920) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = hs_acres / 1e6)) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis") + 
  theme_void() + 
  labs(fill = "Cumulative acres claimed\n(millions), 1920") + 
  theme(legend.position = "bottom")

p2<-dat %>% 
  filter(year == 1920) %>% 
  mutate(hs_pop = hs_claims / tot_pop) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = hs_pop)) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis") + 
  theme_void() +
  labs(fill = "Claims per resident\n1920")  +
  theme(legend.position = "bottom")

grid.arrange(p1, p2, nrow = 1)
```

@@

```{r ts-homestead-claims, fig.cap = "Cumulative Homestead act claims by state and year"}
ggplot(hs %>% 
         filter(year < 1940),
       aes(x = year,
           y = hs_claims,
           group = state)) + 
  geom_line(alpha = 0.5) + 
  geom_text(data = hs %>% 
              filter(year == 1935),
            aes(x = year,
                y = hs_claims,
                label = state),
            position = position_nudge(x = 8),
            alpha = 0.7,
            size = 2.5) + 
facet_wrap(~region)  + 
  labs(y = "Number of claims",
       x = "Year",
       subtitle = "") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

NATIVE LANDHOLDINGS OVER TIME ARE DIRECTLY RELATED TO HOMESTEAD CLAIMS, MECHANISM OF DISPOSESSION.

PULL AGENTS DATA FOR LANDHOLDINGS?

### Urbanization and European immigration

```{r immig-map, fig.cap = "White immigrant population in cities 1920"}
dat %>% 
  filter(year == 1920) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = urban_immig_pop)) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis",
                        trans = "log10") + 
  theme_void() + 
  labs(fill = "Population\n(note logarithmic scale)")
```

```{r urban-ts, fig.cap = "Urban white immigrant population by state 1880-1920"}
p1<-ggplot(dat%>% 
             filter(!(is.na(region))),
           aes(x = year, y = urban_immig_pop / tot_pop,
               group = state)) + 
  geom_line(alpha = 0.5) + 
  labs(subtitle = "Proportion of population white immigrant urban",
       y = "Proportion",
       x = "Year") + 
  facet_wrap(~region)

p2<-ggplot(dat%>% 
             filter(!(is.na(region))),
           aes(x = year, y = urban_immig_pop,
               group = state)) + 
  geom_line(alpha = 0.5) + 
  labs(subtitle = "White immigrant urban population",
       y = "Count",
       x = "Year")+ 
  facet_wrap(~region)

@@

MAYBE ALSO NOTE WHITE IMMIG POP IN FRONTIER STATES, DIFFERENT REGIMES FOR INCORPORATION, DIFFERENT RACIALIZATION PROCESSES FOR NORTH / W EURO IMMIGS VS SOUTH / EAST / IRISH?

## Geographic variation in social and carceral state capacities 1880 - 1920

what are states doing? how does it vary across places and over time?

### Employment in core roles for the emergent welfare and carceral states

@@

### capacity for therapeutic coercion: kids in institutional group quarters

@@

```{r gq-box, fig.cap = "Children in institutional group quarters 1930"}
dat %>% 
  filter(year == 1930) %>% 
  mutate(across(child_inst_correctional:child_inst_poor_disabled, 
          ~ .x/children * 1e3)) %>% 
  pivot_longer(cols = child_inst_correctional:child_inst_poor_disabled,
               names_to = "type",
               values_to = "pop") %>% 
  ggplot(aes(x = type, y = pop, 
             color = region)) + 
  geom_boxplot()
```

@@

<!-- employment trajectories look kinda log-linear. almost certainly non-linear, so need a curve on trajectories regardless. log probably solves it well enough.  -->

<!-- q1: salience of settler expropriotion -> state formation -->

<!-- four models, each employment outcome + children in poverty institutions -->

<!-- start with lm for log(rates) -->

<!-- trouble fitting models for sparse data (e.g. social workers, children in gq), but works fine and in expected directions for teachers, police.  -->

<!-- switch to brms for estimation on tricky ones -->

<!-- ```{r} -->
<!-- ### FIT PROBLEM WAS WITH EARLY YEARS NEAR ZERO AND LOW VARIANCE -->
<!-- model_dat <- dat %>%  -->
<!--   filter(year>=1880) %>%  -->
<!--   mutate(police_rate = (police + 1) / tot_pop * 1e5, -->
<!--          socwork_rate = (socwork + 1) / tot_pop * 1e5, -->
<!--          teacher_rate = (teachers + 1) / tot_pop * 1e5, -->
<!--          child_inst_rate = (child_inst_poor_disabled + 0.1) /  -->
<!--            children * 1e5, -->
<!--          prop_blk = (blk + 1) / tot_pop, -->
<!--          prop_urban_immig = urban_immig_pop / tot_pop, -->
<!--          hsclaim_prop = hs_claims / tot_pop, -->
<!--          year = scale(year)) -->

<!-- m_police<-brm(log(police_rate) ~  -->
<!--                  scale(hsclaim_prop) + -->
<!--                  scale(prop_urban_immig) + -->
<!--                  scale(prop_blk) +  -->
<!--                  scale(log(blk_sharecrop + 1)) +  -->
<!--                  year +  -->
<!--                  (1 + year|state), -->
<!--                data = model_dat) -->

<!-- m_teacher<-brm(log(teacher_rate) ~  -->
<!--                  scale(hsclaim_prop) + -->
<!--                  scale(prop_urban_immig) + -->
<!--                  scale(prop_blk) +  -->
<!--                  scale(log(blk_sharecrop + 1)) +  -->
<!--                  year +  -->
<!--                  (1 + year|state), -->
<!--                data = model_dat) -->

<!-- m_socwork<-brm(log(socwork_rate) ~  -->
<!--                  scale(hsclaim_prop) + -->
<!--                  scale(prop_urban_immig) + -->
<!--                  scale(prop_blk) +  -->
<!--                  scale(log(blk_sharecrop + 1)) +  -->
<!--                  year +  -->
<!--                  (1 + year|state), -->
<!--                data = model_dat) -->

<!-- sc_teacher<-lmer(log(teacher_rate) ~  -->
<!--               scale(hsclaim_prop) + -->
<!--               year +  -->
<!--               (1 + year|state), -->
<!--             data = model_dat, -->
<!--             control=lmerControl(optimizer="nloptwrap", -->
<!--                                 optCtrl=list(maxfun=2e5))) -->

<!-- sc_socwork<-brm(log(socwork_rate) ~  -->
<!--               scale(hsclaim_prop) + -->
<!--               year +  -->
<!--               (1 + year|state), -->
<!--             data = model_dat) -->

<!-- sc_gq<-brm(log(child_inst_rate) ~  -->
<!--               scale(hsclaim_prop) + -->
<!--               year +  -->
<!--               (1 + year|state), -->
<!--             data = model_dat) -->

<!-- jc_police<-lmer(log(police_rate) ~  -->
<!--               scale(prop_blk) + -->
<!--               year +  -->
<!--               (1 + year|state), -->
<!--             data = model_dat, -->
<!--             control=lmerControl(optimizer="nloptwrap", -->
<!--                                 optCtrl=list(maxfun=2e5))) -->

<!-- jc_teacher<-lmer(log(teacher_rate) ~  -->
<!--               scale(prop_blk) + -->
<!--               year +  -->
<!--               (1 + year|state), -->
<!--             data = model_dat, -->
<!--             control=lmerControl(optimizer="nloptwrap", -->
<!--                                 optCtrl=list(maxfun=2e5))) -->

<!-- jc_socwork<-lmer(log(socwork_rate) ~  -->
<!--               scale(prop_blk) + -->
<!--               year +  -->
<!--               (1 + year|state), -->
<!--             data = model_dat, -->
<!--             control=lmerControl(optimizer="nloptwrap", -->
<!--                                 optCtrl=list(maxfun=2e5))) -->

<!-- jc_gq<-lmer(log(child_inst_rate) ~  -->
<!--               scale(prop_blk) + -->
<!--               year +  -->
<!--               (1 + year|state), -->
<!--             data = model_dat, -->
<!--             control=lmerControl(optimizer="nloptwrap", -->
<!--                                 optCtrl=list(maxfun=2e5))) -->
<!-- ``` -->

<!-- This all looks fine. How do I want to visualize this?  -->

<!-- we'll eventually do trajectory visuals for each regime type - what is the expected trajectory under different regimes, where do we end up in 1930? Think about how to handle to intercepts / slopes. What would a zero mean? pulling out everything not directly in national trend and regime measures. is that what I'm after? I guess I really do just want the regime-specific trend lines, in which case E[rate | regime] should zero out the individual idioscyncracies. it's oversimplified, but that should be ok. Just be super clear on what the iterpretation is on prediction. it's not a hypothetical state, its the mean expectation for the slope of the regime var.    -->

<!-- model is  -->

<!-- \[\hat{Y} = \alpha + \beta \textrm{Regime} + \delta year + \epsilon\] -->

<!-- Here we are very much interested just in $\beta$, the rest are nuissance parameters. that works. can vis e[state type] under diffferent REGIME sets.  -->

<!-- Q: are there places with competing regime sets? CA, WA, others? where we need a multivariate regime model?  -->

<!-- ```{r} -->
<!-- new_dat<-data.frame(year =  c(1.3, 1.3, 1.3), -->
<!--                     hsclaim_prop = c(0, 0.05, 0.15)) -->
<!-- t<-predict(sc_police,  -->
<!--            newdata = new_dat, -->
<!--            re.form = ~0, -->
<!--            type = "link") -->

<!-- t1<-predict(sc_teacher,  -->
<!--            newdata = new_dat, -->
<!--            re.form = ~0, -->
<!--            type = "link") -->

<!-- exp(t1) -->

<!-- ``` -->
